---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# labtools

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/joe-m-shaw/labtools/graph/badge.svg)](https://app.codecov.io/gh/joe-m-shaw/labtools)
<!-- badges: end -->

`labtools` is a package for exploratory data analysis at the North West Genomic 
Laboratory Hub.
`labtools` aims to be a centralised, public repository of functions to
help explore genetic data in reproducibly and at scale in lab development projects.

## Installation

You can install the development version of labtools from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("joe-m-shaw/labtools")
```

## Information Governance

**No patient identifiable information should be included in this repository.**

If you spot some, please let me know.

For the purpose of testing functions, I have used generic examples for 
worksheet (WS123456) and lab number (12345678) values.

Where examples of patient names are required, I have used character names from novels by Leo Tolstoy (Anna Karenina, Pierre Bezukhov etc).

## Example

You can use `labtools` to extract sample identifiers stored in filenames, which is useful when doing data projects involving multiple sample files.

```{r}

library(labtools)

filename <- "WS123456_12345678a"

extract_worksheet(filename)

extract_labno(filename)

extract_suffix(filename)

```

A more detailed example is given in the "Reading files with labtools" vignette.
