---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# labno2

<!-- badges: start -->
<!-- badges: end -->

`labno2` is a package that makes it easier to work with sample
identifiers used at the North West Genomic Laboratory Hub.

# What does it mean?

Each sample is given a DNA “lab number” which is stored as “labno” in
sample databases.

## Installation

You can install the development version of labno2 from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("joe-m-shaw/labno2")
```

## Information Governance

**No patient identifiable information should be included in this repository.**

If you spot some, please let me know.

For the purpose of testing functions, I have used generic examples for 
worksheet (WS123456) and lab number (12345678) values.

Where examples of patient names are required, I have used character names from novels by Leo Tolstoy (Anna Karenina, Pierre Bezukhov etc).

## Use of Claude Code

I have used Claude Code to;

- Suggest code improvements

- Generate initial function structure

- Generate fake data for testing functions

## Example

You can use `labno2` to extract sample identifiers stored in filenames, which is useful when doing data projects involving multiple sample files.

```{r}

library(labno2)

filename <- "WS123456_12345678a"

```

Using `labno2`, you can quickly extract the worksheet number that the sample was tested on

```{r}

extract_worksheet(filename)

```

You can extract the lab number, and its suffix (if present).
Samples which are tested multiple times per worksheet will have suffixes of a, b, c etc.

```{r}

extract_labno(filename)

extract_suffix(filename)

```

`labno` also includes the `mutate_ids` function to extract all this information in one go from a dataframe column.

``` {r}

data <- data.frame(
  "filename" = c(filename),
  "gene_result" = c(1)
)

mutate_ids(data)

```
