---
title: "Exploratory data analysis with labtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reading-files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(labtools)
library(readr)
library(purrr)
library(ggplot2)
library(dplyr)
```

# The problem

There are multiple files with genetic results in, and the sample identifiers are only
in the filenames.

```{r files}

files <- list.files(path = "data/",
           pattern = ".*.csv",
           full.names = TRUE)

files

```

Each file consists of results for different genes.

For the purpose of this example, I have randomly generated a "score" for each gene. 
This is not real clinical data.

```{r example-file, echo=FALSE}

file1 <- read_csv(file = files[1],
                  show_col_types = FALSE)

file1

```

I want to get all the results into one dataframe, along with all the relevant identifiers.

# The solution

I can do this for a single file using the `read_csv_with_filename` function from `labtools`.

```{r}

read_csv_with_filename(files[1])

```


Using `map` from `purrr` I can then iterate this function over the file list and 
bind the data together.

```{r read-data}

all_data <- files |> 
  map(\(files) read_csv_with_filename(files)) |> 
  list_rbind() |> 
  mutate_ids() |> 
  mutate(gene_number = parse_number(gene))

```

This gives me the data in a single dataframe with each row annotated with the
correct sample identifiers.

```{r data, echo = FALSE}

knitr::kable(all_data)

```

Then I can explore this collated data using `ggplot2`.

```{r plots, echo = FALSE, width = 6}

gene_plot <- ggplot(all_data, aes(x = gene_number, y = score)) +
  geom_point(shape = 21, size = 2, aes(fill = suffix)) +
  theme_bw()+
  scale_x_continuous(limits = c(1, 10),
                     breaks = seq(1, 10, by = 1)) +
  facet_wrap(~labno) +
  labs(x = "Gene", y = "Score",
       title = "Collated file results",
       subtitle = "Facetted by lab number",
       fill = "Suffix")

gene_plot
  
```

# The rationale

`labtools` is designed to help me do exploratory data analysis on multiple files quickly
and reproducibly.
It is built around `tidyverse` packages including `stringr`, `readr` and `dplyr`.

`labtools` aims to include functionality which is shared across multiple projects that
I do. 
This is shown in the above example: a common problem is collating data from files
which include identifiers in the filenames with a consistent format for
worksheet, lab number and suffix.

# What about extracting patient names?

There is currently no function for extracting patient names in `labtools` for the following reasons:

- **Patient names are not always included in filenames.** If data is analysed on the cloud,
then patient names are removed from filenames, which means you can have files with 
a mixture of patient names and no patient names depending on how they were processed.

- **Not all samples are from people.** Some samples are control samples, such as mixtures 
of different DNA samples or cell lines. These samples may be given a "name" or they
may not, and the name may or may not follow the standard format of "forename surname".
Example: "HorizonPositiveControl".

- **Some samples may be for research.** Patient samples collected for research will 
be anonymised, and the research identifier may be included in place of the sample
name. The research identifier may also include numbers or other non-letter characters.
Example: "study0001"

These reasons mean that designing a regular expression to identify a patient name
in a filename is complicated.
One option is to use the position of the name in the string relative to more 
consistent identifiers, like lab number and worksheet number, but this is then 
inflexible when the order of identifiers changes.
